---
title: "HIV status disclosure and ART adherence among ART-experienced adolescents and young adults in AFRICOS"
output: html_document
author: "Nicole Dear"
date: "`r Sys.Date()`"
---


```{r setup, include = FALSE}

## SETUP -----

# global settings
rstudioapi::writeRStudioPreference("data_viewer_max_columns", 1000L)  # view more columns

# load packages
# install.packages(c("readxl", "ggplot2", "dplyr", "tidyr", "expss", "ggpubr", "stringr", "tidyverse", "lubridate",
#                   "table1", "kableExtra", "lmtest", "sandwich", "k5", "binom", "glm2"))

library(readxl)
library(haven)
library(ggplot2)
library(dplyr)
library(tidyr)
library(textclean)
library(ggpubr)
library(stringr)
library(tidyverse)
library(lubridate)
library(table1)
library(kableExtra)
library(data.table)
library(lmtest)
library(sandwich)
library(lme4)

# set working directory
setwd("G:/DCAC/DCAC_PEPFAR/RV329/Data In/30124 freeze")

# load data
hivstat0 <- read_sas("hivstat0.sas7bdat")
names(hivstat0) <- tolower(names(hivstat0))

arvcode0 <- read_sas("arvcode0.sas7bdat")
names(arvcode0) <- tolower(names(arvcode0))

subjq_dm0 <- read_sas("subjq_dm0.sas7bdat")
names(subjq_dm0) <- tolower(names(subjq_dm0))

subjq_a0 <- read_sas("subjq_a0.sas7bdat")
names(subjq_a0) <- tolower(names(subjq_a0))

subjq_c0 <- read_sas("subjq_c0.sas7bdat")
names(subjq_c0) <- tolower(names(subjq_c0))

v_load0 <- read_sas("v_load0.sas7bdat")
names(v_load0) <- tolower(names(v_load0))

```


### Project Description

Adolescents and young adults represent an increasing share of people living with HIV (PLWH) globally. In 2022, 1.65 (1.18-2.19) million of an estimated 39.0 million total PLWH were age 10-19 years.1 Adolescents and young adults with HIV face multifaceted challenges to long-term treatment success, particularly if HIV was acquired at an early age. Barriers include limited availability of pediatric-formulated regimens, accumulation of drug resistance mutations, treatment fatigue, and socio-behavioral factors such as lack of social support, violence, disclosure concerns and stigma that may interfere with optimal ART adherence.2–5 Strategies to improve ART adherence are critical to attaining the UNAIDS third 95 among all age groups of PLWH.  

Prior evidence suggests that HIV status disclosure is associated with improved HIV outcomes, including ART adherence, enhanced engagement in care, better immune recovery and viral suppression as well as higher likelihood of condom use and increased social support.6,7 A recent meta-analysis found that children and adolescents who were aware of their HIV status had improved adherence to ART compared to those who were not aware of their status.8 However, data on disclosure to others (ex., peers, caregivers) is limited.  

The objective of this study is to characterize HIV status disclosure among adolescents and young adults in AFRICOS and assess the relationship between status disclosure and ART adherence as well as assess whether route of HIV transmission (vertical vs non-vertical) may moderate the relationship between status disclosure and ART adherence.  


```{r, include = FALSE}

## DATA CLEANING -----

hivstat1 <- hivstat0 %>%
  select(subjid, visit, visitdt, gender, age, dobdtn, hivflag, hivstat, progid, diagdtn, art_sdtn, dur_art, dur_hiv) %>% 
  filter(visit<90) %>% 
  distinct(subjid, visit, .keep_all = TRUE)

# test <- hivstat1 %>% 
#   group_by(subjid) %>%
#   mutate(v1_date = visitdt[1]) # first row by subjid

# check for dups by subjid and visit
# duptest1 <- hivstat1 %>% 
#   group_by(subjid, visit) %>% 
#   mutate(dupe = n()>1)
# table(duptest1$dupe)

subjq_dm1 <- subjq_dm0 %>%
  select(subjid, visitdt, visit, educat, eductxt, employed, enrl_ovc, provide, hhincome, hcurrtyp, hcurrtxt, 
         hholdur, howfrkm, food, mealnum, mealcut) %>% 
  filter(visit<90) %>% 
  # convert unk/no response values to missing
  mutate(enrl_ovc = na_if(enrl_ovc, -1)) %>%
  mutate(enrl_ovc = na_if(enrl_ovc, 8)) %>%
  # remove participants with duplicate visits with wrong date
  filter(!(subjid=="D01-0045" & visit==18 & visitdt=='2023-01-25')) %>%
  filter(!(subjid=="D01-0555" & visit==3 & visitdt=='2023-03-01')) %>% 
  distinct(subjid, visit, .keep_all = TRUE)
  
# check for dups by subjid and visit
# duptest2 <- subjq_dm1 %>% 
#   group_by(subjid, visit) %>% 
#   mutate(dupe = n()>1)
# table(duptest2$dupe)

subjq_a1 <- subjq_a0 %>%
  select(subjid, visit, visitdt, sequence, hivinfct, hivintxt, hivinf, takearv, pillnum, pillnr, missarv, lsmsarv, msarv_a,
         msarv_b, msarv_c, msarv_d, msarv_e, msarv_f, msarv_g, msarv_h, msarv_i, msarv_j, msarv_k, msarv_l, msarv_m, msarv_n, 
         msarv_o, msarv_p, msarv_z, msarvtxt, msarnone, arvcost, acurrtyp, acurrtxt, arvsupp, hivfood, hvfd_a, hvfd_b, hvfd_z, 
         hvfdtxt, hvste_a, hvste_b, hvste_c, hvste_z, hvstetxt, hvsteno, hvstd_a, hvstd_b, hvstd_z, hvstdtxt, hvstdno, whiv_a, 
         whiv1a, whiv_b, whiv2b, whiv_c, whiv3c, whiv_d, whiv4d, whiv_e, whiv5e, whiv_f, whiv6f, whiv_g, whiv7g, whiv_h, 
         whiv8h, whiv_i, whiv9i, whiv_j, whiv10j, whiv_z, whiv11z, whivtxt, whivnone, sghivfrq, mtct, oftkarv, gjobarv, 
         msarv_q, comgrp_a, comgrp_b, comgrp_c, comgrp_d, comgrp_e, comgrp_z, comgrptx, comgrpno, hvsti_a, hvsti_b, hvsti_c, 
         hvsti_d, hvsti_e, hvsti_f, hvsti_g, hvsti_h, hvsti_i, hvsti_j, hvsti_k, hvsti_l, hvsti_m, hvsti_n, hvsti_z, unwlove, 
         whiv_k, whiv11k, whiv_l, whiv12l, whiv_m, whiv13m, whiv_n, whiv14n, attend_a, attend_b, attend_c, attend_d, attend_z,
         othattxt) %>% 
  filter(visit<90 & sequence==1) %>%
  # convert unk/no response values to missing
  mutate(missarv = na_if(missarv, 7)) %>%
  mutate(unwlove = na_if(unwlove, 99)) %>%
  # remove participants with duplicate visits with wrong date
  filter(!(subjid=="D01-0045" & visit==18 & visitdt=='2023-01-25')) %>%
  filter(!(subjid=="D01-0371" & visit==15 & visitdt=='2022-02-21')) %>% 
  filter(!(subjid=="D02-0010" & visit==10 & visitdt=='2023-03-22')) %>% 
  distinct(subjid, visit, .keep_all = TRUE)

# check for dups by subjid and visit
# duptest3 <- subjq_a1 %>% 
#   group_by(subjid, visit) %>% 
#   mutate(dupe = n()>1)
# table(duptest3$dupe)

subjq_c1 <- subjq_c0 %>%
  select(subjid, visit, visitdt, sequence, alcohol, alcohnum, alcdrnum, alcohsex, alcohcut, cigaret, cigarnum, drug, p_hit,
         p_abu, p_threat, p_refuse, p_forced, p_othv, pothtx, g_hit, g_abu, g_threat, g_refuse, g_forced, g_othvix, gothtx,
         violnh_a, violnh_b, violnh_c, violnh_d, violnh_e, violnh_f, violnh_g, violnh_z, violhtxt, viohnone, violnp_a, 
         violnp_b, violnp_c, violnp_d, violnp_e, violnp_f, violnp_g, violnp_z, violptxt, viopnone, harm, harmsxor, 
         sexchild, sexforce, afraid) %>% 
  filter(visit<90) %>%
  # convert unk/no response values to missing
  mutate(alcohol = na_if(alcohol, 8)) %>%
  mutate(cigaret = na_if(cigaret, -1)) %>%
  mutate(drug = na_if(drug, -1)) %>%
  # remove participants with duplicate visits with wrong date
  filter(!(subjid=="D01-0045" & visit==18 & visitdt=='2023-01-25')) %>%
  filter(!(subjid=="D02-0010" & visit==10 & visitdt=='2023-03-22')) %>% 
  distinct(subjid, visit, .keep_all = TRUE)

# check for dups by subjid and visit
# duptest4 <- subjq_c1 %>% 
#   group_by(subjid, visit) %>% 
#   mutate(dupe = n()>1)
# table(duptest4$dupe)

v_load1 <- v_load0 %>%
  select(subjid, visit, visitdt, sequence, drawper, drawdt, vl_circ, vlcopy, nodetect, vl) %>%
  # include only rows with draw at study visit
  filter(drawper==1) %>%
  # remove duplicate rows
  # (for D02-0124, this row with visit=2 is a dup of visit 7
  # and for B01-0033, this row is a dup where the other visit 7 row has non-missing vlcopy)
  filter(!(subjid=="B01-0033" & visit==7 & is.na(vlcopy))) %>%
  filter(!(subjid=="D02-0124" & visitdt=="2022-08-19" & visit==2)) %>%
  # remove rows that have same subjid, visit, and vlcopy
  distinct(subjid, visit, vlcopy, .keep_all = TRUE) %>%
  # variable to check if visit values are repeated within a subjid
  group_by(subjid, visit) %>%
  mutate(dupflag = if_else(row_number()>1, 1, 0)) %>%
  # remove rows from same visit with sequence greater than 1
  filter(!(dupflag==1 & sequence!=1)) %>% 
  # clean vl variable
  mutate(vl_clean = case_when(nodetect==1 & is.na(vlcopy) ~ "undetectable",
                              vl_circ==1 ~ paste0("<", vlcopy),
                              vl_circ==2 | is.na(vl_circ) ~ as.character(vlcopy),
                              vl_circ==3 ~ paste0(">", vlcopy))) %>% 
  mutate(vl1000 = case_when(vl<1000 ~ 0,
                            vl>=1000 ~ 1,
                            .default = NA)) %>% 
  mutate(vls1000 = case_when(vl<1000 ~ 1,
                            vl>=1000 ~ 0,
                            .default = NA))


## MERGE DATA -----

data0 <- hivstat1 %>%
  left_join(subjq_dm1, by = c("subjid", "visit")) %>% 
  left_join(subjq_a1, by = c("subjid", "visit")) %>% 
  left_join(subjq_c1, by = c("subjid", "visit")) %>% 
  left_join(v_load1, by = c("subjid", "visit"))

```


```{r, include = FALSE}

## RECODE FREE TEXT RESPONSES -----

# code free text responses for route of HIV transmission
data1 <- data0 %>% 
  mutate(hivintxt_new = strip(hivintxt, lower.case = FALSE, apostrophe.remove=TRUE))

freetext <-as.data.frame(table(data1$hivintxt_new))

# define patterns to search for
mtct_text <- c('AT BIRTH', 'BORN WITH', 'BREAST MILK', 'BREST FEEDING', 'DURING DELIVERY', 'MOTHER TO CHILD', 'MOTHER TOCHILD', 
            'FROM THE PARENTS', 'PMTCT', 'MTCT', 'VERRTICAL TRANSMISSION', 'VERTICAL')
unk_text <- c('CAN NOT RECALL', 'CANNOT IDENTIFY HOW HE ACQUIRED', 'NOT SURE', 'DO NOT KNOW', 'DOES NOT KNOW', 'DOESNT KNOW',
              'DONT KNOW', 'DONT REMEMBER', 'NOT KNOW', 'NOT KNOWN', 'NOT KNOW', 'CANNOT TELL HOW SHE WAS INFECTED', 'UNKNOWN')

data1$flag_mtct <- ifelse(grepl(paste(mtct_text, collapse='|'), data1$hivintxt), 1, 0)
data1$flag_unk <- ifelse(grepl(paste(unk_text, collapse='|'), data1$hivintxt), 1, 0)


# code free text responses for disclosure to 'other'
freetext2 <-as.data.frame(table(data1$whivtxt))

# define patterns to search for
txt_a <- c('CO-WIFE', 'CO-WIVES', 'CO WIFE', 'HUSBAND', 'PARTNER', 'SPOUSE', 'FANCE', 'FIANCE', 'GILFRIEND', 'GIRL FRIEND', 
           'GIRLFRIEND', 'WIFE', 'SEX PARTNER', 'WIFI')
txt_b <- c('FATHER', 'MOTHER', 'GUARDIAN', 'IN-LAWS', 'INLAW', 'MZAZI', 'PARENT', 'WALEZI')
txt_c <- c('BROTHER', 'SISTER', 'PACHA', 'SHEMEJI')
txt_d <- c('DAUGHTER', 'SON')
txt_colleague <- c('CO-WORKER', 'COLLEAGUE', 'COLLEGUE', 'EMPLOYEE', 'WORKMATE')
txt_e <- c('GRAND MOTHER', 'GRAND PARENT', 'GRANDMOTHER', 'GRANDPARENT')
txt_f <- c('COUSIN', 'PASTOR AND AUNT')
txt_g <- c('NIECE')
txt_h <- c('FELLOW ORPHANS', 'FELLOW TEACHERS', 'FRIEND', 'OTHER PRISONERS', 'PRISONMATE', 'TRAVEL COMPANION')
txt_j <- c('HEAD OF CHURCH', 'CHURCH LEADERS', 'CHURCH MEMBERS', 'CHURCHMATES', 'CHURCH LEADER', 'PSTOR', 'PASTOR', 
           'ISLAMIC LEADER', 'MOSQUE LEADER', 'MZEE WA KANISA', 'WAUMINI KANISANI', 'WAZEE WA KANISA')
txt_k <- c('EMPLOYER', 'EMPLOYMENT', 'BOSS', 'HEAD OF HIS DEPARTMENT', 'HEAD OF HIS WORK', 'HUMAN RESOURCE', 'SUPERVISOR',
           'MANAGER', 'NEW INCHARGE')
txt_l <- c('DESKMATE')
txt_m <- c('SCHOOL LECTURERS', 'SCHOOL MATRON', 'TEACHER')
txt_n <- c('GRANDCHILD', 'GRAND CHILD', 'GRAND CHILDREN', 'GRAND DAUGHERS', 'GRAND DAUGHTER', 'GRAND SON', 'GRANDSON',
           'GRANS SON', 'MJUKUU', 'WAJUKUU')
txt_neighbor <- c('NEIGBOUR', 'NEIGHBOR', 'NEIGHBOUR', 'NEIGHBOUURS', 'JIRANI', 'MY NEIGHBOUR', 'NEIGBOUR',
                  'NEIGHBOR', 'NEIGHBOUS', 'VILLAGEMATES')

data1$flag_a <- ifelse(grepl(paste(txt_a, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_b <- ifelse(grepl(paste(txt_b, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_c <- ifelse(grepl(paste(txt_c, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_d <- ifelse(grepl(paste(txt_d, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_colleague <- ifelse(grepl(paste(txt_colleague, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_e <- ifelse(grepl(paste(txt_e, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_f <- ifelse(grepl(paste(txt_f, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_g <- ifelse(grepl(paste(txt_g, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_h <- ifelse(grepl(paste(txt_h, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_j <- ifelse(grepl(paste(txt_j, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_k <- ifelse(grepl(paste(txt_k, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_l <- ifelse(grepl(paste(txt_l, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_m <- ifelse(grepl(paste(txt_m, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_n <- ifelse(grepl(paste(txt_n, collapse='|'), data1$whivtxt), 1, 0)
data1$flag_neighbor <- ifelse(grepl(paste(txt_neighbor, collapse='|'), data1$whivtxt), 1, 0)

# code free text responses for 'other' education
educat1 <- c("1", "ONE", "2", "TWO", "PRIMARY")
educat3 <- c("3", "FOUR", "SEVEN", "8", "EIGHT", "SECONDARY")
educat5 <- c("COLLAGE", "COLLEGE", "UNIVERSITY")
educat6 <- c("POST GRADUATE")

data1$flag_educat1 <- ifelse(grepl(paste(educat1, collapse='|'), data1$eductxt), 1, 0)
data1$flag_educat3 <- ifelse(grepl(paste(educat3, collapse='|'), data1$eductxt), 1, 0)
data1$flag_educat5 <- ifelse(grepl(paste(educat5, collapse='|'), data1$eductxt), 1, 0)
data1$flag_educat6 <- ifelse(grepl(paste(educat6, collapse='|'), data1$eductxt), 1, 0)

test <- data1 %>%
  # select(subjid, visit, visitdt.x, hivflag, hivinfct, hivintxt, flag_mtct, flag_unk)
  # select(subjid, visit, visitdt.x, hivflag, whivtxt, flag_a, flag_b, flag_c, flag_d, flag_colleague, flag_e, flag_f, flag_g, 
  # flag_h, flag_j, flag_k, flag_l, flag_m, flag_n, flag_neighbor)
  select(subjid, visit, visitdt.x, hivflag, educat, eductxt, flag_educat1, flag_educat3, flag_educat5, flag_educat6)

# create new variables and convert to factor
data2 <- data1 %>%
  
  # vertical vs non-vertical transmission
  mutate(vertical = case_when(hivinfct==8 | flag_mtct==1 ~ 1,
                                hivinfct==1 | hivinfct==2 | hivinfct==3 | hivinfct==4 | hivinfct==5 | hivinfct==6 | 
                                  hivinfct==7 | (hivinfct==90 & flag_mtct!=1 & flag_unk!=1) ~ 0,
                                hivinfct==97 | hivinfct==98 | flag_unk==1 ~ NA,
                                .default = 0)) %>%
  mutate(vertical = factor(vertical,
                    levels = c(0,1),
                    labels = c("Not vertical transmission", "Vertical transmission"))) %>%
  
  # disclosure variables
  mutate(disclose_spouse = case_when(whiv_a==1 | flag_a==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_parent = case_when(whiv_b==1 | flag_b==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_sibling = case_when(whiv_c==1 | flag_c==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_child = case_when(whiv_d==1 | flag_d==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_colleague = case_when(flag_colleague==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_grandparent = case_when(whiv_e==1 | flag_e==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_extendadult = case_when(whiv_f==1 | flag_f==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_extendadol = case_when(whiv_g==1 | flag_g==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_friend = case_when(whiv_h==1 | flag_h==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_church = case_when(whiv_j==1 | flag_j==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_employer = case_when(whiv_k==1 | flag_k==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_schoolmate = case_when(whiv_l==1 | flag_l==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_teacher = case_when(whiv_m==1 | flag_m==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_grandchild = case_when(whiv_n==1 | flag_n==1 ~ 1, 
                                .default = 0)) %>% 
  mutate(disclose_neighbor = case_when(flag_neighbor==1 ~ 1, 
                                .default = 0)) %>%
  mutate(disclose_any = if_else(disclose_spouse==1 | disclose_parent==1 | disclose_sibling==1 | disclose_child==1 |
                                  disclose_colleague==1 | disclose_grandparent==1 | disclose_extendadult==1 |
                                  disclose_extendadol==1 | disclose_friend==1 | disclose_church==1 | disclose_employer==1 |
                                  disclose_schoolmate==1 | disclose_teacher==1 | disclose_grandchild==1 | 
                                  disclose_neighbor==1, 1, 0)) %>%
  mutate(disclose_immedfam = if_else(disclose_spouse==1 | disclose_parent==1 | disclose_sibling==1 | 
                                       disclose_child==1, 1, 0)) %>%

  # demographics
  mutate(country = case_when(progid==1 ~ 1,
                             progid==2 | progid==3 ~ 2,
                             progid==4 ~ 3,
                             progid==5 ~ 4,
                             .default = NA)) %>%
  mutate(country = factor(country,
                          levels = c(1,2,3,4),
                          labels = c("Uganda", "Kenya", "Tanzania", "Nigeria"))) %>% 
  mutate(progid = factor(progid,
                   levels = c(1,2,3,4,5),
                   labels = c("Kayunga, Uganda", "South Rift Valley, Kenya", "Kisumu West, Kenya", 
                              "Mbeya, Tanzania", "Abuja & Lagos Nigeria"))) %>%
  mutate(hholdur = factor(hholdur,
                          levels = c(1,2,3),
                          labels = c("Urban", "Peri-urban", "Rural"))) %>%
  mutate(gender = factor(gender,
                   levels = c(1,2),
                   labels = c("Male", "Female"))) %>%
  mutate(hivflag = factor(hivflag,
                    levels = c(1,2),
                    labels = c("PLWH", "PLWoH"))) %>% 
  mutate(hivstat = factor(hivstat,
                    levels = c(1,2),
                    labels = c("PLWH", "PLWoH"))) %>%
  mutate(education = case_when(educat==0 | educat==1 | (educat==90 & flag_educat1==1) ~ 0,
                               educat==2 | educat==3 | (educat==90 & flag_educat3==1) ~ 1,
                               educat>3 | (educat==90 & flag_educat5==1) | (educat==90 & flag_educat6==1) ~ 2,
                               .default = NA)) %>% 
  mutate(education = factor(education,
                          levels = c(0,1,2),
                          labels = c("None or some primary", "Primary or some secondary", "Secondary and above"))) %>%
  
  # ART adherence & treatment support
  mutate(missdose = case_when(missarv==0 ~ 0,
                              missarv>=1 ~ 1,
                              .default = NA)) %>%
  mutate(nomissdose = case_when(missarv==0 ~ 1,
                              missarv>=1 ~ 0,
                              .default = NA)) %>%
  # mutate(missdose = factor(missdose,
  #                          levels = c(0,1),
  #                          labels = c("No missed doses", "Missed 1 or more doses"))) %>%
  mutate(missarv = factor(missarv,
                          levels = c(0,1,2,3,4),
                          labels = c("None", "1-2 days", "3-5 days", "6-10 days", "More than 10 days"))) %>%
  mutate(arvsupp = factor(arvsupp,
                           levels = c(0,1),
                           labels = c("No treatment supporter", "Has treatment supporter"))) %>%
  mutate(hivsupportgroup = case_when(sghivfrq==0 ~ 0,
                                     sghivfrq==1 ~ 1,
                                     sghivfrq==2 ~ 2,
                                     sghivfrq==3 | sghivfrq==4 ~ 3,
                                     .default = NA)) %>%
  mutate(hivsupportgroup = factor(hivsupportgroup,
                           levels = c(0,1,2,3),
                           labels = c("Not at all", "Less than once a month", "Once a month", "More than once a month"))) %>%
  mutate(enrl_ovc = factor(enrl_ovc,
                      levels = c(0,1,5),
                      labels = c("No", "Yes", "Over age 18 (NA)"))) %>%
  
  # food insecurity
  mutate(mealnum_c = case_when(mealnum<3 ~ 0,
                               mealnum==3 ~ 1,
                               mealnum>3 ~ 2,
                               .default = NA)) %>% 
  mutate(mealcut_c = case_when(mealcut==0 ~ 0,
                               mealcut>=1 ~ 1,
                               .default = NA)) %>% 
  mutate(foodinsecure = case_when(food==1 & (mealnum_c==1 | mealnum_c==2) ~ 0,
                                  food==0 | mealnum_c==0 ~ 1,
                                  .default = NA)) %>% 
  mutate(mealnum_c = factor(mealnum_c,
                            levels = c(0,1,2),
                            labels = c("<3 meals/day", "3 meals/day", ">3 meals/day"))) %>% 
  mutate(mealcut_c = factor(mealcut_c,
                            levels = c(0,1),
                            labels = c("No meals cut/reduced", "1+ meals cut/reduced"))) %>%
  mutate(foodinsecure = factor(foodinsecure,
                            levels = c(0,1),
                            labels = c("Enough Food and 3+ meals/day", "Not Enough Food or <3 meals/day"))) %>%
  
  # substance use
  mutate(alcohol = factor(alcohol,
                          levels = c(0,1),
                          labels = c("No", "Yes"))) %>%
  mutate(cigaret = factor(cigaret,
                          levels = c(0,1),
                          labels = c("No", "Yes"))) %>% 
  mutate(drug = factor(drug,
                          levels = c(0,1),
                          labels = c("No", "Yes"))) %>%
  
  # stigma and discrimination
  mutate(discrim_disrupt_care = case_when(hvstd_a==1 | hvstd_b==1 ~ 1,
                                  .default = 0)) %>% 
  mutate(discrim_disrupt_care = factor(discrim_disrupt_care,
                                       levels = c(0,1),
                                       labels = c("No", "Yes"))) %>%
  mutate(unwlove = factor(unwlove,
                          levels = c(0,1,2,3),
                          labels = c("Never", "Rarely", "Sometimes", "Often"))) %>%
  group_by(subjid) %>%
  fill(c(hivflag, gender, diagdtn, art_sdtn, progid, dobdtn, hholdur, vertical, disclose_spouse, disclose_parent,
         disclose_sibling, disclose_child, disclose_colleague, disclose_grandparent, disclose_extendadult, 
         disclose_extendadol, disclose_friend, disclose_church, disclose_employer, disclose_schoolmate, disclose_teacher,
         disclose_grandchild, disclose_neighbor, disclose_any, disclose_immedfam), .direction = "downup") %>%
  ungroup()

# fill in any missing agev using calculated age
i <- is.na(data2$age)
data2$age[i] <- with(data2, (data2$visitdt.x[i] - data2$dobdtn[i]) / 365.25)

# fill in duration since HIV dx and duration on ART if missing
j <- is.na(data2$dur_hiv)
data2$dur_hiv[j] <- difftime(data2$visitdt.x[j], data2$diagdtn[j], units = "days") / 365.25

k <- is.na(data2$dur_art)
data2$dur_art[k] <- difftime(data2$visitdt.x[k], data2$art_sdtn[k], units = "days") / 365.25

```


```{r, echo = FALSE}

## ENROLLMENT SUMMARY -----

# plot enrollment by HIV status and site, overall
s1 <- data2 %>%
  filter(visit==1) %>%
  group_by(hivflag, progid) %>% 
  summarise(Freq = n())

p1 <- ggplot(s1, aes(x=progid, y=Freq, fill=hivflag)) +
  geom_bar(stat="identity", position=position_dodge()) +
  labs(x = "", y = "Number of participants") +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
        geom_text(aes(label=Freq), vjust=1.6, color="white", 
                  position = position_dodge(0.9), size=3.5) +
        scale_fill_brewer(palette="Paired")

print(p1)

# plot enrollment by HIV status and site, youth cohort only
s2 <- data2 %>%
  filter(visit==1 & age<25) %>%
  group_by(hivflag, progid) %>% 
  summarise(Freq = n())

p2 <- ggplot(s2, aes(x=progid, y=Freq, fill=hivflag)) +
  geom_bar(stat="identity", position=position_dodge()) +
  labs(x = "", y = "Number of participants") +
  theme(legend.position = "bottom",
        legend.title = element_blank()) +
        geom_text(aes(label=Freq), vjust=1.6, color="white", 
                  position = position_dodge(0.9), size=3.5) +
        scale_fill_brewer(palette="Paired")

print(p2)

```


```{r, echo = FALSE}

# DEFINE ANALYTIC POPULATION -----

# PLWH and PLWoH enrolled
data2 %>%
  group_by(subjid) %>%
  count(hivflag) %>%
  group_by(hivflag) %>%
  count()

# PLWH and PLWoH enrolled, youth and young adults (under 25 years)
data2 %>%
  filter(age<25) %>% 
  group_by(subjid) %>%
  count(hivflag) %>%
  group_by(hivflag) %>%
  count()

# PLWH on ART at most recent visit, youth and young adults (under 25 years)
pop0 <- data2 %>%
  filter(hivflag=="PLWH" & takearv==1 & age<25) %>%
  group_by(subjid) %>%
  slice_tail(n = 1)

# pop0 <- data2 %>%
#   filter(hivflag=="PLWH" & takearv==1 & age<25)

table(pop0$hivflag)

# complete cases
pop_cc0 <- pop0[complete.cases(pop0[, c('gender', 'age', 'progid', 'vertical', 'missdose', 'vl1000', 'disclose_spouse', 'disclose_parent', 'disclose_sibling', 'disclose_child', 'disclose_colleague', 'disclose_grandparent', 'disclose_extendadult', 'disclose_extendadol', 'disclose_friend', 'disclose_church', 'disclose_employer', 'disclose_schoolmate', 'disclose_teacher', 'disclose_grandchild', 'disclose_neighbor', 'disclose_any')]),]

table(pop_cc0$hivflag)

# setwd("G:/DCAC/DCAC_PEPFAR/RV329/Analyses/Youth/Disclosure_adherence")
# write.xlsx(pop_cc0, file = "data_set_clean.xlsx")

```


```{r, echo = FALSE}

## SUMMARIZE DISCLOSURE VARIABLES -----

s3 <- pop_cc0 %>%
  group_by(hivflag) %>%
  summarise(anyone = sum(disclose_any),
            immediatefam = sum(disclose_immedfam),
            spouse = sum(disclose_spouse),
            parent = sum(disclose_parent),
            sibling = sum(disclose_sibling),
            child = sum(disclose_child),
            colleague = sum(disclose_colleague),
            grandparent = sum(disclose_grandparent),
            extadult = sum(disclose_extendadult),
            extadol = sum(disclose_extendadol),
            friend = sum(disclose_friend),
            church = sum(disclose_church),
            friend = sum(disclose_friend),
            employer = sum(disclose_employer),
            schoolmate = sum(disclose_schoolmate),
            teacher = sum(disclose_teacher),
            grandchild = sum(disclose_grandchild),
            neighbor = sum(disclose_neighbor),
            n = n()) %>% 
  select(anyone, immediatefam, spouse, parent, sibling, child, colleague, grandparent, extadult, extadol, friend, church, 
         employer, schoolmate, teacher, grandchild, neighbor)

s3_t <- s3 %>%
  rownames_to_column() %>% 
  pivot_longer(!rowname, names_to = "disclosed_to", values_to = "col2") %>% 
  pivot_wider(names_from = "rowname", values_from = "col2") %>% 
  rename("freq" = "1")

# calculate percentages
s3_t$percent <- round(100*(s3_t$freq/698),2)

# create labels
s3_t$label <- paste(s3_t$freq, " (", round(s3_t$percent, digits = 1), "%)", sep = "")

# create factor version of disclosed_to variable
s3_t$disclosed_to_f <- NA
s3_t$disclosed_to_f[s3_t$disclosed_to=="anyone"] <- 1
s3_t$disclosed_to_f[s3_t$disclosed_to=="child"] <- 2
s3_t$disclosed_to_f[s3_t$disclosed_to=="church"] <- 3
s3_t$disclosed_to_f[s3_t$disclosed_to=="colleague"] <- 4
s3_t$disclosed_to_f[s3_t$disclosed_to=="employer"] <- 5
s3_t$disclosed_to_f[s3_t$disclosed_to=="extadol"] <- 6
s3_t$disclosed_to_f[s3_t$disclosed_to=="extadult"] <- 7
s3_t$disclosed_to_f[s3_t$disclosed_to=="friend"] <- 8
s3_t$disclosed_to_f[s3_t$disclosed_to=="grandchild"] <- 9
s3_t$disclosed_to_f[s3_t$disclosed_to=="grandparent"] <- 10
s3_t$disclosed_to_f[s3_t$disclosed_to=="immediatefam"] <- 11
s3_t$disclosed_to_f[s3_t$disclosed_to=="neighbor"] <- 12
s3_t$disclosed_to_f[s3_t$disclosed_to=="parent"] <- 13
s3_t$disclosed_to_f[s3_t$disclosed_to=="schoolmate"] <- 14
s3_t$disclosed_to_f[s3_t$disclosed_to=="sibling"] <- 15
s3_t$disclosed_to_f[s3_t$disclosed_to=="spouse"] <- 16
s3_t$disclosed_to_f[s3_t$disclosed_to=="teacher"] <- 17

s3_t <- s3_t %>% 
  mutate(disclosed_to_f = factor(disclosed_to_f,
                          levels = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17),
                          labels = c("Anyone", "Child (son/daughter)", "Church member", "Colleague", "Employer",
                                     "Extended adolescent family member", "Extended adult family member", "Friend/peer",
                                     "Grandchild", "Grandparent", "Immediate family member", "Neighbor", "Parent", 
                                     "Schoolmate", "Sibling", "Spouse/partner", "Teacher")))

# plot
bar <- s3_t %>%
  arrange(freq) %>%
  mutate(disclosed_to_f=factor(disclosed_to_f, levels=disclosed_to_f)) %>%
  ggplot(aes(x = disclosed_to_f, y = percent)) + 
  geom_bar(stat = "identity", fill="royalblue") +
  coord_flip() +
  geom_text(aes(label = label), size = 3.5, hjust = -0.2) +
  xlab("Individual dislosed to") +
  ylab("Percent") +
  lims(y = c(0, 50))

plot(bar)

```


```{r, echo = FALSE}

## DESCRIPTIVE ANALYSES -----

pop_cc <- pop_cc0 %>% 
  mutate(disclose_spouse = factor(disclose_spouse,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_parent = factor(disclose_parent,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_sibling = factor(disclose_sibling,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_child = factor(disclose_child,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_colleague = factor(disclose_colleague,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_grandparent = factor(disclose_grandparent,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_extendadult = factor(disclose_extendadult,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_extendadol = factor(disclose_extendadol,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_friend = factor(disclose_friend,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_church = factor(disclose_church,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_employer = factor(disclose_employer,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_schoolmate = factor(disclose_schoolmate,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_teacher = factor(disclose_teacher,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_grandchild = factor(disclose_grandchild,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_neighbor = factor(disclose_neighbor,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>%
  mutate(disclose_any = factor(disclose_any,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>% 
  mutate(disclose_immedfam = factor(disclose_immedfam,
                    levels = c(0,1),
                    labels = c("Not disclosed", "Disclosed"))) %>% 
  mutate(missdose = factor(missdose,
                            levels = c(0,1),
                            labels = c("No missed doses", "Missed 1 or more doses"))) %>%
  mutate(vl1000 = factor(vl1000,
                            levels = c(0,1),
                            labels = c("VL<1000 c/ml", "VL>=1000 c/ml")))

# table 1
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}


# t1 <- table1(~ gender + age + progid + hholdur + education + vertical + dur_art + missarv + missdose +
#                arvsupp + hivsupportgroup + enrl_ovc + mealnum_c + mealcut_c + foodinsecure + alcohol + cigaret + drug +
#                discrim_disrupt_care + unwlove + disclose_spouse + disclose_parent + disclose_sibling + disclose_child +
#                disclose_colleague + disclose_grandparent + disclose_extendadult + disclose_extendadol + disclose_friend +
#                disclose_church + disclose_employer + disclose_schoolmate + disclose_teacher + disclose_grandchild +
#                disclose_neighbor + disclose_any, pop0)
# t1flex(t1)

t2 <- table1(~ vertical + missdose + vl1000 + disclose_spouse + disclose_parent + disclose_sibling + disclose_child +
               disclose_colleague + disclose_grandparent + disclose_extendadult + disclose_extendadol + disclose_friend +
               disclose_church + disclose_employer + disclose_schoolmate + disclose_teacher + disclose_any +
               disclose_immedfam | gender, pop_cc, overall=c(right="Total"), extra.col=list(`P-value`=pvalue))
t1flex(t2)

t3 <- table1(~ disclose_spouse + disclose_parent + disclose_sibling + disclose_child +
               disclose_colleague + disclose_grandparent + disclose_extendadult + disclose_extendadol + disclose_friend +
               disclose_church + disclose_employer + disclose_schoolmate + disclose_teacher + disclose_any +
               disclose_immedfam | vertical, pop_cc, overall=c(right="Total"),
             extra.col=list(`P-value`=pvalue))
t1flex(t3)

# note: a greater % of those with HIV via vertical transmission have a treatment supporter -- kindof interesting..

# t.test(pop_cc2$age ~ pop_cc2$vertical, alternative = "two.sided", var.equal = FALSE)

```


```{r, echo = FALSE}

## MULTIVARIABLE CROSS-SECTIONAL ANALYSES -----

# change reference levels
# pop_cc2$gender <- relevel(pop_cc2$gender, ref = "Male")


## DISCLOSURE --> ART ADHERENCE (1: missed doses ART)

# robust poisson and apply the sandwich error estimator
model <- glm(missdose ~ disclose_immedfam, data=pop_cc0, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()

# robust poisson and apply the sandwich error estimator
model <- glm(missdose ~ disclose_immedfam + gender + age + progid, data=pop_cc0, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()


## DISCLOSURE --> VIRAL NONSUPPRESSION (1: VL>=1000 c/ml)

# robust poisson and apply the sandwich error estimator
model <- glm(vl1000 ~ disclose_immedfam, data=pop_cc0, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()

# robust poisson and apply the sandwich error estimator
model <- glm(vl1000 ~ disclose_immedfam + gender + age + progid + missdose, data=pop_cc0, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()


## TEST INTERACTION TERMS

# robust poisson and apply the sandwich error estimator
model <- glm(missdose ~ disclose_immedfam*vertical, data=pop_cc0, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()


## Note: I am seeing effect modification by sex...

# robust poisson and apply the sandwich error estimator
model <- glm(missdose ~ disclose_immedfam*gender + age + progid, data=pop_cc0, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()


# robust poisson and apply the sandwich error estimator
model <- glm(vl1000 ~ disclose_immedfam*gender + age + progid + missdose, data=pop_cc0, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()


## Stratify models by sex

pop_cc_f <- pop_cc0 %>% filter(gender=="Female")
pop_cc_m <- pop_cc0 %>% filter(gender=="Male")

## ART adherence

# robust poisson and apply the sandwich error estimator
model <- glm(nomissdose ~ disclose_immedfam + age + progid, data=pop_cc_f, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()

# robust poisson and apply the sandwich error estimator
model <- glm(nomissdose ~ disclose_immedfam + age + progid, data=pop_cc_m, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()


## Viral suppression

# robust poisson and apply the sandwich error estimator
model <- glm(vls1000 ~ disclose_immedfam + age + progid + missdose, data=pop_cc_f, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()

# robust poisson and apply the sandwich error estimator
model <- glm(vls1000 ~ disclose_immedfam + age + progid + missdose, data=pop_cc_m, family = poisson(link="log"))

# pull out robust coefficients
model_out <- coeftest(model, vcov = sandwich)

# calculate CIs
model_out_ci <- round(cbind(exp(cbind(RR = model_out[,1], 
                              LCI = model_out[,1] + qnorm(0.05/2)*model_out[,2],
                              UCI = model_out[,1] - qnorm(0.05/2)*model_out[,2])),
                              P = model_out[,4]),2)

# output the table
model_out_ci %>%
  kable() %>%
  kable_styling()

```


```{r, echo = FALSE}

## MULTIVARIABLE LONGITUDINAL ANALYSES -----

## DISCLOSURE --> ART ADHERENCE (1: missed doses ART)

# estimate the model and store results in m (this uses lme4, currently not able to load Matrix package)
m1<- glmer(missdose ~ disclose_immedfam + (1 | subjid), data = pop_cc0, family = binomial,
           control = glmerControl(optimizer = "bobyqa"), nAGQ = 10)

# print the mod results without correlations among fixed effects
print(m1, corr = FALSE)

# table of estimates with 95% CI
se <- sqrt(diag(vcov(m1)))
tab <- cbind(Est = fixef(m1), LL = fixef(m1) - 1.96 * se, UL = fixef(m1) + 1.96 * se)
exp(tab)

```